Fast, Scalable Phrase-Based SMT Decoding

### 2016-05-03 acl.2016/acl2016.tex
changes

The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimzations, including cube-pruning. It is widely used for MT research and commercial use.

@@

The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{https://twitter.com/ApacheJoshua/status/342022794097340416}).

@@

A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once. Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis. This process is called \emph{hypothesis expansion} and starts with a hypothesis that has translated no source word and ends with a completed hypotheses that have translated all source words. The highest-scoring completed hypothesis, according to the model score, is returned as  most probable translation, $\hat{t} $.

In the phrase-based model, each rule translates a contiguous sequence of source words but successive translation options do not have to be adjacent on the source side, depending on the distortion limit. The target output is constructed strictly left-to-right from the target side of successive translation rules.

@@

The first three feature functions frequently trained on data and require the feature to read the model from files. The other feature functions are simple and do not leave much room for optimization. The language model consumes a large part of decoding time, and it is for this reason that it has already been heavily optimized. %We shall therefore investigate the first two feature functions for optimization.

@@

For any realistic sized phrase-based model to be used in an online situation, memory and loading time constraints requires us to use load-on-demand phrase-table implementations. Moses contains a number of such implementations with different performance characterics, we show the time take to decode 800,000 sentences for the fastest two in Figure~\ref{fig:moses-phrase-tables-time}.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding time with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.

@@

There are four main optimizations that we shall investigate. Firstly, the search creates and destroy a large number of hypothesis objects in memory which puts a burden on the operating system. We shall optimize the search algorithm to use memory pools and object pools, replacing the operating system's general purpose memory management with our own application-aware management.

@@

@@

Thirdly, we will look at for optimizating the phrase-table, specifically in relation to the Probing implementation, focusing on two techniques. One, we will investigate caching of often used translation rules. The default caching framework save the most recently looked up rules. However, this caching mechanism actually makes decoding slower, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}
We shall explore a simpler caching mechanism that creates a static cache at the start of decoding.

@@

>>>>

The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.

@@

The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{\scriptsize https://twitter.com/ApacheJoshua/status/342022794097340416}).

@@

A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once. Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis. This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that have translated all source words. The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.

In the phrase-based model, each rule translates a contiguous sequence of source words. Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit. The target output is constructed strictly left-to-right from the target side from the series of translation rules.

@@

Of these feature functions, the last four are simple and do not leave much room for optimization. As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\newcite{Heafield-kenlm} and ~\newcite{yasuhara-EtAl:2013:EMNLP}.

Therefore, we will focus on optimizing the functionality of the phrase-table and lexicalized re-ordering model.

@@

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding time with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.

@@

There are four main optimizations that we shall investigate.

\subsection{Efficient Memory Allocation}

Firstly, the search algorithm creates and destroy a large number of hypothesis and related objects. The memory allocation required puts a burden on the operating system due to the need to synchronize  memory access. This problem worsens with multi-threaded applications such as Moses.

Libraries such as tcmalloc~\cite{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.

We shall seek to improve decoding speed by replacing the operating system's general purpose memory management with our own custom memory management scheme. In time-critical parts of the decoder, memory will be allocated from a memory pool.

A memory pool is a large block of memory that has been pre-allocated by the operating system from which smaller blocks can be give out to the application. We will use a thread-specific memory pools to avoid locking contention during memory access. Our memory pool will be dynamic, growing in size when required but never reducing in size.

To further increase memory management speed, objects are not deleted, instead the memory pool is simply reset and reused for each input sentence. This can result in high memory usage so object queues are available for high-churn objects which allows the decoder to re-use portion of the memory pool before it is reset. The object queues are LIFO so that recently accessed memory are used, increasing CPU cache hits.

@@

\subsection{Stack Configurations}

@@

\subsection{Phrase-Table Optimizations}

Thirdly, we will look at for optimizating the phrase-table, specifically in relation to the Probing implementation, focusing on two techniques.

One, the default translation rule caching mechanism in Moses saves the most recently used rules. However, this is a complex mechanism which require locking when timestamping datastructures and periodic clearing of old rules in order in order to reduce memory consumption. It actually makes decoding slower, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}

We shall explore a simpler caching mechanism that creates a static cache at the start of decoding of the most likely translation rules to be used.

@@

\subsection{Lexicalized Reordering Model Optimizations}
